Neural models for computer-assisted musical orchestration: a preliminary study

### 2020-05-12 paper/latex/paper.tex
Revise data section

\section{Data generation}

\subsection{Sources}
We used the TinySOL database to create our input data. TinySOL is a subset of the Studio On Line (SOL) database created by IRCAM \textbf{Reference needed}. TinySOL contains 1,529 samples from 12 instruments. The instruments come from different orchestral families: strings, woodwinds, and brass. Notes are played using the ordinario style and with three different dynamics for each pair of instrument and pitch : \textit{pp}, \textit{mf} and \textit{ff}. The exact list of samples is shown in Tab.\ref{tab_pitch}.

Since the goal was to train the model to detect multiple instruments playing simultaneously, the samples used for training were generated by mixing a random combination of instruments and pitches. More precisely, given the number $N$ of instruments playing at the same time, the data was generated as follows:

\begin{itemize}
\item An orchestra of $N$ instruments is chosen in advance (it is the same for all samples).
\item For each instrument in the orchestra, a random note is picked among the corresponding available samples in TinySOL.
\item For each chosen note, a random dynamic is picked.
\item The samples are added and normalized by the number of instruments.
\item The Mel spectrogram of the mixture is then computed and used as input of the model.
\end{itemize}

The choice of Mel spectrograms as input features of the model is very commmon in music information \textbf{\{Insert reference\}} and can be considered to be an accurate representation of the way humans perceive sound and music.\\

Samples had a sample rate of $44100Hz$ and were all padded or chopped to be exactly 4 seconds long. The Mel spectrograms were then generated using an FFT hop length of 2048 samples (e.g. the window of each FFT was $46ms$ wide), and a number of Mel bins equal to 128. Therefore, all the features fed to the model were matrices of size $128\times 345$. We used Librosa package in Python to compute the features, and more details on the exact computations can be found in \textbf{\{Reference to Librosa\}} The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10ms$ could be perceived by the human ear.

@@

\subsection{Data augmentation}

@@

@@

\begin{equation}
M_N(\hat{X})_i = \left\{\begin{array}{ll}
1 \text{ if } i \in I_N(\hat{X})\\
0 \text{ otherwise}
\end{array}\right.
\label{NMax}
\end{equation}
and $I_N(\hat{X})$ is the set of the indices containing the $N$ first maximums of the vector $\hat{X}$.\\

More specifically, the function $M_N(\hat{X})$ takes as an input a vector of probabilities (values between 0 and 1) and outputs a vector where only the positions of the $N$ first maxima are set to 1. This new vector would be the orchestration of $N$ instruments given by the model. Thus, the function $f$ simply outputs the proportion of the estimated orchestration that matches the expected one.\\

Different experiments were made by varying the number of instruments in the orchestra $N$ but always matching this number with the number of instruments in each mixture sample. Tab. \ref{exp_tab} shows the orchestra for each different value of $N$ used in the experiments.\\

@@

\section{Evaluation and Conclusions}
- the approach seems promising for orchestration
- many things used in Orchidea are not implemented here (symbolic constraints, sparsity,...)
- CNN seems better for timbre
- ResNet seems better for pitch (what are timbre and pitch??)

@@

\textbf{(Alejandro, add how there were certain instruments that performed poorly. Add how ResNet was awful for Flute and always output Flute C4)}

>>>>

\section{Data}

\subsection{Data Source and Generation}
We used the TinySOL database to create our input data. TinySOL is a subset of the Studio On Line (SOL) database created by IRCAM \textbf{Reference needed}. TinySOL contains 1,529 samples from 12 instruments. The instruments come from different orchestral families: strings, woodwinds, and brass. Each sample is one instrument playing a single note in the ordinario style, with one of three dynamics: \textit{pp}, \textit{mf}, or \textit{ff}. The instruments and ranges over which they were recorded is summarized in Tab.\ref{tab_pitch}.

We generated all the data we used to train our models. For a chosen $N$, each input to our model is a combination of $N$ TinySOL samples. An orchestra of $N$ instruments is chosen and the data is generated in the following way:

\begin{itemize}
\item For each instrument in the orchestra, a random TinySOL sample for that instrument is chosen. This leads to a random combination of both notes and dynamics.
\item The chosen samples are combined to be played simultaneously and normalized by the number of instruments.
\item The mel spectrogram of the mixture is computed to be used as the features input to the model.
\end{itemize}

The choice of using the mel spectrogram as input features of the model is very commmon in music information \textbf{\{Insert reference\}} and can be considered to be an accurate representation of the way humans perceive sound and music.

Samples had a sample rate of $44100$Hz and were all padded or chopped to be exactly 4 seconds long. The Mel spectrograms were then generated using an FFT hop length of 2048 samples (e.g. the window of each FFT was $46$ms wide), and a number of Mel bins equal to 128. Therefore, all the features fed to the model were matrices of size $128\times 345$. We used Librosa package in Python to compute the features, and more details on the exact computations can be found in \textbf{\{Reference to Librosa\}} The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10$ms could be perceived by the human ear.

@@

\subsection{Data Augmentation}

@@

- explain why we generated our data in the way we did
Since our goal is to train a model to detect multiple instruments and pitches playing simultaneously, the data used for training is generated by mixing a random combination of instruments and pitches.

@@

\begin{equation}
M_N(\hat{X})_i = \left\{\begin{array}{ll}
1 \text{ if } i \in I_N(\hat{X})\\
0 \text{ otherwise}
\end{array}\right.
\label{NMax}
\end{equation}
and $I_N(\hat{X})$ is the set of the indices containing the $N$ first maximums of the vector $\hat{X}$.

More specifically, the function $M_N(\hat{X})$ takes as an input a vector of probabilities (values between 0 and 1) and outputs a vector where only the positions of the $N$ first maxima are set to 1. This new vector would be the orchestration of $N$ instruments given by the model. Thus, the function $f$ simply outputs the proportion of the estimated orchestration that matches the expected one.

Different experiments were made by varying the number of instruments in the orchestra $N$ but always matching this number with the number of instruments in each mixture sample. Tab. \ref{exp_tab} shows the orchestra for each different value of $N$ used in the experiments.

\textbf{(Alejandro, add how there were certain instruments that performed poorly and some that were great. Add how ResNet was awful for Flute and always output Flute C7)}

@@

\section{Conclusion}
- the approach seems promising for orchestration
- many things used in Orchidea are not implemented here (symbolic constraints, sparsity,...)
- CNN seems better for timbre
- ResNet seems better for pitch (what are timbre and pitch??)

@@
